**First high-frequency time series of carbonate chemistry in the Arctic**

### 2022-03-31 awipev-CO2_presentation.Rmd
final

---
title: "**First high-frequency time series of carbonate chemistry in the Arctic**" 
author: "Jean-Pierre Gattuso^1,2^, Samir Alliouane^1^ & Philipp Fisher^3^"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document:
    # code_folding: show
    #css: acid.css
    fig_caption: no
    #fig_width: 2
    #fig_height: 2
    highlight: default    
    number_sections: no
    #theme: yeti
    theme: cerulean
    toc: no
    toc_float: no
    fontsize: 32pt
#language: Australian
#css: acid.css
## mainfont: PT Serif  
## monofont: PT Mono
#csl: frontiers.csl
## sansfont: PT Sans
#bibliography: ArcticKelp.bib
## papersize: A4
---
<!-- <style type="text/css"> -->
<!-- .main-container { -->
<!--   max-width: 100% !important; -->
<!--   margin: auto; -->
<!-- } -->
<!-- </style> -->
<style type="text/css">
.main-container {
  max-width: 100% !important;
  margin: auto;
}
h1.title {
  font-size: 50px;
}
body{
  font-size: 24pt;
}
h1,h2,h3,h4,h5,h6{
  font-size: 32pt;
}
.blackbox {
  padding: 1em;
  background: LightYellow;
  color: black;
  border: 0px solid orange;
  border-radius: 10px;
}
.center {
  text-align: center;
}
</style>

@@

@@

- Arctic region:  one of the "**reasons for concern**" of the IPCC.
- Arctic Ocean: less than 1% of ocean water but receives 11% of the global runoff; 7-10% of the global burial of organic carbon
- **SST** increased twice as fast as other regions; projections of up to +5.5 °C in 2100 
- Projected decrease in surface **pH** by 2100 larger than elsewhere (-0.45 vs -0.06 to -0.32 units)
- **Undersaturation** of surface water with respect to aragonite ﬁrst reported in 2008 in the Canada Basin; due to the decrease of salinity and degradation of organic matter
- **No** fewer observations than elsewhere for carbonate variables: 11.6% of the SOCAT pCO<sub>2</sub> records and 9.1% of the GLODAP C<sub>T</sub> records are from the Arctic Ocean which is only about 4.3% of the global ocean surface
<!-- - Coastal data relatively well represented in both products (25% and 10% of the SOCAT and SOCAT Arctic data, respectively) -->
- **Uneven distribution**: 71.5% of the SOCAT pCO<sub>2</sub> data and 72% of the GLODAP C<sub>T</sub> data collected Jun-Sep). Few to very few data Dec-Mar.
- Critical to document the changes but no high frequency multi-year time-series
:::

@@

# **Availability of data**
:::: {style="display: flex;"}
::: {class='col-md-4'}
Many challenges:

@@

Nevertheless:

@@

- **Positive aspects**: 
  - **First** multi-year, high-frequency time series of the carbonate system in the Arctic 
  - Excellent support from AWIPEV staff until 2020
- **Logistical difficulties**:
  - Winter sampling essential
  - Challenging environment; divers needed to sort out issues
  - Difficult and costly to gather a team and organize field trip at short notice
  - New regulation for using HgCl<sub>2</sub> and transporting poisoned samples

>>>>

---
title: "**First high-frequency time series of carbonate chemistry in the Arctic**" 
author: "Jean-Pierre Gattuso^1,2^, Samir Alliouane^1^ & Philipp Fisher^3^"
date: '`r format(Sys.Date(), "%d %B %Y")`'
output:
  html_document:
    # code_folding: show
    #css: acid.css
    fig_caption: no
    #fig_width: 2
    #fig_height: 2
    highlight: default    
    number_sections: no
    #theme: yeti
    theme: cerulean
    toc: no
    toc_float: no
    fontsize: 32pt
#language: Australian
#css: acid.css
## mainfont: PT Serif  
## monofont: PT Mono
#csl: frontiers.csl
## sansfont: PT Sans
#bibliography: ArcticKelp.bib
## papersize: A4
---
<!-- <style type="text/css"> -->
<!-- .main-container { -->
<!--   max-width: 100% !important; -->
<!--   margin: auto; -->
<!-- } -->
<!-- </style> -->
<style type="text/css">
.main-container {
  max-width: 100% !important;
  margin: auto;
}
h1.title {
  font-size: 50px;
}
body{
  font-size: 24pt;
}
h1,h2,h3,h4,h5,h6{
  font-size: 32pt;
}
h2,h3,h4,h5,h6{
  font-size: 28pt;
}
.blackbox {
  padding: 1em;
  background: LightYellow;
  color: black;
  border: 0px solid orange;
  border-radius: 10px;
}
.center {
  text-align: center;
}
</style>

@@

<br><br><br><br><br><br><br><br><br><br><br><br><br>

@@

- Arctic region:  one of the "**reasons for concern**" of the IPCC.
- Arctic Ocean: less than 1% of ocean water but receives 11% of the global runoff; 7-10% of the global burial of organic carbon
- **SST** increased twice as fast as other regions; projections of up to +5.5 °C in 2100 
- Projected decrease in surface **pH** by 2100 larger than elsewhere (-0.45 vs -0.06 to -0.32 units)
- **Undersaturation** of surface water with respect to aragonite ﬁrst reported in 2008 in the Canada Basin; due to the decrease of salinity and degradation of organic matter
- **No** fewer observations than elsewhere for carbonate variables: 11.6% of the SOCAT pCO<sub>2</sub> records and 9.1% of the GLODAP C<sub>T</sub> records are from the Arctic Ocean which is only about 4.3% of the global ocean surface
<!-- - Coastal data relatively well represented in both products (25% and 10% of the SOCAT and SOCAT Arctic data, respectively) -->
- **Uneven time distribution**: 71.5% of the SOCAT pCO<sub>2</sub> data and 72% of the GLODAP C<sub>T</sub> data collected Jun-Sep. Few to very few data Dec-Mar.
- Critical to document the changes but no high frequency multi-year time-series
:::

@@

# **Availability of data**
:::: {style="display: flex;"}
::: {class='col-md-4'}
## Many challenges:

@@

## Nevertheless:

@@

- **Positive aspects**: 
  - First multi-year, high-frequency time series of the carbonate system in the Arctic 
  - Excellent support from AWIPEV staff until 2020
- **Logistical difficulties**:
  - Winter sampling essential
  - Challenging environment; divers needed to sort out issues
  - Difficult and costly to gather a team and organize field trip on short notice
  - New regulation for the use HgCl<sub>2</sub> and transportation of poisoned samples
